---
title: "Virus Infection Dynamics Calculations"
author: "Elijah I. Christenson and Qingyang Zhang"
date: "2023-05-19"
output:
  html_document: default
  md_document: default
  pdf_document: default    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

 renv::init()
 

## Calculates I~SC~, as well as AUC and other necessary values to achieve I~SC~
   This loads necessary packages and finds N~asymptote~ for both control and infected

```{r}
install.packages("dplyr",repos = "http://cran.us.r-project.org")
install.packages("ggplot2",repos = "http://cran.us.r-project.org")
install.packages("readxl",repos = "http://cran.us.r-project.org")
install.packages("DescTools",repos = "http://cran.us.r-project.org")


library("dplyr")
library("ggplot2")
library("readxl") 
library("DescTools")

```

## Read in your data to the R environment from excel using Import Dataset function
 Data should be in the following format in excel:
 
 hpi=Hours post infection, abs=absorbance at 600nm

time.hpi | ctrl.abs | virus1.abs | ctrl.sd | virus1.sd 

```{r}
testing = read_xlsx("testing.xlsx")
time0 = data.frame(time.hpi=0,   
                   ctrl.abs=0,   
                   ctrl.sd=0,
                   virus1.abs=0,
                   virus1.sd=0,
                   virus2.abs=0,
                   virus2.sd=0)
testing = bind_rows(time0, testing)
```

```{r}
#the following finds asymptotes
time=testing[c("time.hpi")] #change to specify column
con_abs=testing[c("ctrl.abs")] #change to specify column
Nasymcon <- max(con_abs) #picks max value from specified column
con_inf=testing[c("virus1.abs")] #change to specify column
Nasyminf <- max(con_inf) #pick max value from specified column
```

## To calculate PI~max~, I~SC~, PI~AUC~, and V~R~
Using the equations below from "Quantifying relative virulence", https://doi.org/10.1099/jgv.0.001515
$$PI_{max} = (1- N_{asymptote(infected)}/N_{asymptote(control)}) * 100$$
$$I_{SC} = [PI_{AUC} * PI_{max}]^{1/2}$$
$$PI_{AUC} = (1- AUC_{(infected)}/AUC_{(control)}) * 100$$
$$V_{R} = V_{R} + (I_{SC}[i+1] + I_{SC}[i])*(t[i+1]-t[i])/(t_{n}-t_{1})/2$$
```{r}
PIMAX=(1-(Nasyminf/Nasymcon))*100 #calculates PImax
AUCcon=AUC(testing$time.hpi[1:16], testing$ctrl.abs[1:16])#AUC for control, change ratio as needed, ratio should be [1:total number of time points to include]
AUCinf=AUC(testing$time.hpi[1:16], testing$virus1.abs[1:16])#AUC for infected, change ratio as needed 
PIAUC=(1-(AUCinf/AUCcon))*100 #calculates PIauc
Isc=(PIAUC*PIMAX)^(0.5) #calculates Isc

# To calculate VR, we calculate N.asy.inf, N.asy.ctr, AUC.inf, AUC.ctr, PI.AUC, PI.max and I.sc for each upperbound
time = testing$time.hpi[1:16]
ctr.abs = testing$ctrl.abs[1:16]
inf.abs = testing$virus1.abs[1:16]

Nasyminf.vec = numeric()
Nasyminf.vec[1] = 0
Nasymcon.vec = numeric()
Nasymcon.vec[1] = 0
AUCinf.vec = numeric()
AUCinf.vec[1] = 0
AUCcon.vec = numeric()
AUCcon.vec[1] = 0
PIAUC.vec = numeric()
PIAUC.vec[1] = 0
PIMAX.vec= numeric()
PIMAX.vec[1] = 0 
Isc.vec = numeric()
Isc.vec[1] = 0

for(i in 2:length(time)){
  time.temp = time[1:i]
  ctr.abs.temp = ctr.abs[1:i]
  inf.abs.temp = inf.abs[1:i]
  Nasyminf.vec[i] = max(inf.abs.temp)
  Nasymcon.vec[i] = max(ctr.abs.temp)
  AUCinf.vec[i] = AUC(time.temp, inf.abs.temp)
  AUCcon.vec[i] = AUC(time.temp, ctr.abs.temp)
  PIAUC.vec[i] = (1-AUCinf.vec[i]/AUCcon.vec[i])*100
  PIMAX.vec[i] = (1-Nasyminf.vec[i]/Nasymcon.vec[i])*100
  Isc.vec[i] = sqrt(PIAUC.vec[i]*PIMAX.vec[i])
}

results = data.frame(time,
                     ctr.abs,
                     inf.abs,
                     Nasymcon.vec,
                     Nasyminf.vec,
                     AUCcon.vec,
                     AUCinf.vec,
                     PIAUC.vec,
                     PIMAX.vec,
                     Isc.vec)

results

# Finally we calculate VR 
VR = 0 # initial value
# Note that here t0 is not involved, so t1 = time[2]
for(i in 2:(nrow(results)-1)){
  VR = VR + (Isc.vec[i+1] + Isc.vec[i])*(time[i+1]-time[i])/(time[nrow(results)]-time[2])/2
}

VR

```


### Code to make graph with asymptotes and shading
    This helps to better visualize the I~SC~ and AUC
```{r}
ggplot(data=testing, aes(x=time.hpi)) + 
  
  theme_light(base_size = 10)+
  geom_point(aes(y = ctrl.abs), color = "darkred") + #copy to add more datasets
  geom_point(aes(y = virus1.abs), color="steelblue")+
  geom_errorbar(aes(ymin=ctrl.abs-ctrl.sd, ymax=ctrl.abs+ctrl.sd),
                color="darkred", width=.1)+
  geom_errorbar(aes(ymin=virus1.abs-virus1.sd, ymax=virus1.abs+virus1.sd),
                color="steelblue", width=.1)+ #copy geom_errorbar to add data

  geom_vline(xintercept=which.max(testing$ctrl.abs), linetype= "solid", color= "black", linewidth=0.75)+ 
#this is a line that should intersect at the time point with the highest absorbance value on the control
  
  geom_hline(yintercept= Nasymcon, linetype= "dashed", color= "darkred", linewidth=0.70)+ #Horizontal lines, intersect curve at highest absorbance value
  geom_hline(yintercept= Nasyminf, linetype= "dashed", color= "steelblue", linewidth=0.70)+
  annotate("text", x = 2.2, y = 0.56, label = "N[asymptote]", vjust = -0.5)+ #dashed line labels, adjust to where you want it on graph
  
  ggtitle("testgraph")+ #main title
  
  labs(x="Hours post-infection (hpi)", y="Absorbance (OD 600nm)")+ #axis labels
  geom_ribbon(data = testing %>% filter(time.hpi < 12), ymin=-Inf, 
              aes(ymax=ctrl.abs), fill='darkred', alpha=0.6)+ 
geom_ribbon( data = testing %>% filter(time.hpi < 12), ymin=-Inf, 
             aes(ymax=virus1.abs), fill='steelblue', alpha=0.4) #Fill colors, you will need to change the (time.hpi < intercept+1) argument to one integer higher than your vertical x-intercept line
  
```

renv::snapshot()
